---
title: "Opinion poll analysis"
author: "Varun Khanna"
date: "2 April 2019"
output: github_document
always_allow_html: yes
---

```{r global_options, include=TRUE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=8, fig.path='Figs/',warning = FALSE, message = FALSE)
```

# Indian elections 

The world's largest election (Indian lok sabha elections 2019) with 879 million eligible voters and 8000+ candidates is around the corner, therefore, it apt to analyze the opinion poll data to guess the mood of the electorate. In 2014 nearly half of the Indian voters preferred a party other than the BJP and its rival Congress. Let us see what does opinion polls tell us this time. In this report I have summarized the opinion poll data available in public on the wikipedia page given below: 

#### __Acknowledgement__:  All the data is derived from [the wikipedia page](https://en.wikipedia.org/wiki/Opinion_polling_for_the_2019_Indian_general_election) of Indian general elections 2019.

## Load packages and define the functions

```{r load_packages}
# Load the packages
library("tidyverse")
library("kableExtra")
library("lubridate")
library("knitr")
library("janitor")
library("scales")
library("ggmap")
library("maps")
library("rgdal")
library("maptools")
library("gridExtra")
library("rgeos")

# Function to read the data and clean it up
read_clean <- function(skip, nrows)
{
  data <- read.csv("raw_data.csv", skip = skip, nrows = nrows, header = TRUE, as.is = TRUE )
data <- remove_empty(data, which = "cols")
# For consitency remove the spaces
data$Polling.agency <- gsub(" ","",data$Polling.agency)

return(data)
}

# Function to conver the data into long format
make_long <- function(df){
  df <- df %>% select(., -Lead)
  long_df <- gather(df, key = "party", value = "seats", -c(Date.published, Polling.agency))
 # Replace - sign with 0
  long_df <- long_df %>% mutate_at(vars(3:length(long_df)), funs(str_replace_all(., "-","0")))
 long_df$date <- parse_date_time(long_df$Date.published,orders = c("bdy", "bY"))
 
 long_df$seats <- as.integer(long_df$seats)
 return(long_df)
  }

# Create a blank theme
blank_theme <- theme_minimal()+
theme(
axis.title.x = element_blank(),
axis.text.x=element_blank(),
axis.title.y = element_blank(),
panel.border = element_blank(),
panel.grid=element_blank(),
axis.ticks = element_blank(),
plot.title=element_text(size=14, face="bold")
)
```



## Opinion polling

```{r opinion_poll}
opinion_poll <- read_clean(1,19)

# Look at the data
kable(opinion_poll) %>% kable_styling()

long_opinion_poll <- opinion_poll %>% select(1,2,3,4,5) %>% gather(key = "party", value = "seats", c(-Date.published, -Polling.agency))

long_opinion_poll$date <- parse_date_time(long_opinion_poll$Date.published,orders = c("bdy", "bY"))
long_opinion_poll$party <- factor(long_opinion_poll$party, levels =  c("NDA", "UPA","Others"))

# Plot for all 
long_opinion_poll %>% ggplot(aes( x= Polling.agency, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge()) + scale_fill_manual(values=c('#FFA500', '#228B22', '#1E90FF'), name = "Party", labels = c("NDA", "UPA", "Others")) + theme(axis.text.x = element_text(angle = 90)) + ggtitle("Opinion polls") + facet_wrap(~ reorder(Date.published, date)) + labs(x = "Polling agency", y = "Number of seats")

mean_long_op <- long_opinion_poll %>% group_by(Date.published, party, date) %>% summarise(average_seats = round(mean(seats),0))

# Plot avergae of all opinion polls for the month
ggplot(mean_long_op, aes(x = reorder(Date.published, date), y = average_seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge()) + scale_fill_manual(values=c('#FFA500', '#228B22', '#1E90FF'), name = "Party", labels = c("NDA", "UPA", "Others")) + labs(y = "Average seats", x = "Polling month") + geom_text(aes(label = average_seats), col = "black", size = 3, vjust = 1.5, position = position_dodge(0.9)) + theme(axis.text.x = element_text(angle = 90)) + ggtitle("Average of opinion polls")
```

# Vote share

```{r vote_share}

vote_share <- read_clean(24,6)

# Look at the data
kable(vote_share) %>% kable_styling()

long_vote_share <- vote_share %>% select(1,2,3,4,5) %>% gather(key = "party", value = "vote_share_percent", c(-Date.published, -Polling.agency))

long_vote_share$date <- parse_date_time(long_vote_share$Date.published,orders = c("bdy", "bY"))

long_vote_share$party <- factor(long_vote_share$party, levels =  c("NDA", "UPA","Others"))

ggplot(data = long_vote_share, mapping = aes(x = Polling.agency, y = vote_share_percent, col = party)) + geom_point(aes(size = vote_share_percent)) + facet_wrap(~reorder(Date.published, date)) + scale_color_manual(values=c('#FFA500', '#228B22', '#1E90FF'), name = "Party", labels = c("NDA", "UPA", "Others")) + labs(y = "Vote Percentage", x = "Polling Agency") + theme(axis.text.x = element_text(angle = 90)) + ggtitle("Percentage vote share") + scale_size_discrete(guide=FALSE)
```

# Next PM

```{r nextpm}
next_pm <- read_clean(33,6)

# Look at the data
kable(next_pm) %>% kable_styling()

long_next_pm <- next_pm %>% select(1,2,3,4,5) %>% gather(key = "person", value = "percent", c(-Date.published, -Polling.agency, -Lead))

long_next_pm$date <- parse_date_time(long_next_pm$Date.published,orders = c("bdy", "bY"))

ggplot(long_next_pm, aes(x = reorder(Date.published, date), y = percent, fill = person)) + geom_bar(stat = "identity", position = position_dodge()) + scale_fill_manual(values=c('#FFA500', '#228B22'), name = "Person", labels = c("Narendra Modi", "Rahul Gandhi")) + labs(y = "Percentage in favour", x = "Polling month") + theme(axis.text.x = element_text(angle = 90)) + ggtitle("Voters preference for next PM") + facet_wrap(~Polling.agency)
```

# Narendra Modi (NM) vs Priyanka Gandhi Vadra (PGV)

Do people want to see a contest between NM and PGV?

```{r nm_pgv}
nm_pgv <- read_clean(42,1)

# Look at the data
kable(nm_pgv) %>% kable_styling()

long_nm_pgv <- nm_pgv %>% select(1,2,3,4,5) %>% gather(key = "response", value = "percent", c(-Date.published, -Polling.agency, -Lead))
# Remove percent sign
long_nm_pgv$percent <- as.numeric(gsub("%","",long_nm_pgv$percent))

pie <- ggplot(long_nm_pgv, aes(x = "", y = percent, fill = response)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start = 0)

pie + blank_theme + geom_text(aes(y = percent/3 + c(0, cumsum(percent)[-length(percent)]),label = percent(percent/100)), size=5) + scale_fill_manual(values = c("#999999", "#E69F00"))
```

# Economic performance; compared to previous government

```{r economic_performance}
economics <- read_clean(46,3)
# Look at the data
kable(economics) %>% kable_styling()

long_economics <-  economics %>% select(1,2,3,4,5) %>% gather(key = "response", value = "percent", c(-Date.published, -Polling.agency))
long_economics$percent <- as.numeric(gsub("%","",long_economics$percent))

long_economics$date <- parse_date_time(long_economics$Date.published,orders = c("bdy", "bY"))

ggplot(long_economics, aes(x = "", y = percent, fill = response, group = date)) + geom_bar(width = 1, stat = "identity") + geom_text(aes(label = percent(percent/100), group = date), position = position_stack(vjust = 0.5)) + facet_grid(Polling.agency ~ reorder(Date.published, date)) + coord_polar("y", start = 0) + blank_theme + scale_fill_manual(values=c('#FFA500', '#1E90FF', '#228B22'), name = "Response", labels = c("Better", "Don\'t know", "Worse"))
```

# Key Issues

```{r key_issues}
key_issues <- read_clean(52,1)
# Look at the data
kable(key_issues) %>% kable_styling()

long_key_issues <-  key_issues %>% select(1:6) %>% gather(key = "issues", value = "percent", c(-Date.published, -Polling.agency))

long_key_issues$date <- parse_date_time(long_key_issues$Date.published,orders = c("bdy", "bY"))

ggplot(long_key_issues, aes(x = issues, y = percent)) + geom_bar(stat = "identity", width = 0.5) + scale_x_discrete(labels = c("Corruption","Farmers plight","Price rise","Unemployment")) + labs(x = "Key issues", y = "Percentage of response") + theme_minimal()
```

## Other issues

```{r other_issues}
other_issues <- read_clean(56,11)
other_issues <- other_issues[c(1:3,7,11),]

# remove the percentage sign
other_issues <- other_issues %>% mutate(Yes = as.numeric(str_remove(Yes, "%")), No = as.numeric(str_remove(No, "%")), Don.t.know = as.numeric(str_remove(Don.t.know, "%")), Lead = as.numeric(str_remove(Lead, "%")))

# Calculate the mean of the first three surveys
meanValues <- other_issues[1:3,3:5] %>% apply(., 2, mean) %>% round()

other_issues <- other_issues[c(1,4:5),]
other_issues[1,3] <- as.vector(meanValues[1])
other_issues[1,4] <- as.vector(meanValues[2])
other_issues[1,5] <- as.vector(meanValues[3])


issue <- c("Farm distress", "Jobs enough created","Corruption in Rafale deal")
other_issues$issue <- issue

# Look at the data
kable(other_issues) %>% kable_styling()

long_other_issues <- other_issues %>% select(1:5,7) %>% gather(key = "response", value = "percent", c(-Date.published, -Polling.agency, -issue))

long_other_issues$date <- parse_date_time(long_other_issues$Date.published,orders = c("bdy", "bY"))

ggplot(long_other_issues, aes(x = response, y = percent, fill = issue )) + geom_bar(stat = "identity", position = position_dodge()) + scale_fill_manual(values=c('#FFA500', '#1E90FF', '#228B22'), name = "Response", labels = c("Corruption in Rafale deal", "Farm distress", "Jobs enough created")) + labs(x = "Response", y = "Percentage of respondents") + ggtitle("Other issues")
```

# State wise polling

## Andhra Pradesh
```{r andhra_pradesh}
andhra_pradesh <- read_clean(71,2)
# Look at the data
kable(andhra_pradesh) %>% kable_styling()

long_andhra_pradesh <- make_long(andhra_pradesh)

ggplot(long_andhra_pradesh, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge()) + facet_wrap(~Polling.agency) + scale_fill_manual(values=c('#FFA500', '#FFD700', '#228B22', '#008B8B'), name = "Party", labels = c("NDA", "TDP", "UPA","YSRCP")) + labs(x = "", y = "Number of seats") + ggtitle("Andhra Pradesh (25)")

meanPoll2019 <- data.frame(party = character(),
                  average_seats = double(),
                  state = character(), stringsAsFactors = FALSE)

# Calculate the mean poll for 2019
meanPoll <- filter(long_andhra_pradesh, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise(average_seats = round(mean(seats))) %>% mutate(state = "Andhra Pradesh") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Assam
```{r assam}
assam <- read_clean(80,2)

# Look at the data
kable(assam) %>% kable_styling()

long_assam <- make_long(assam)

ggplot(long_assam, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~Polling.agency) + scale_fill_manual(values = c('#00FF00', '#FFA500', '#228B22'), name = "Party", labels = c("AIUDF","NDA", "UPA")) + labs(x = "", y = "Number of seats") + ggtitle("Assam (14)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_assam, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise(average_seats = round(mean(seats))) %>% mutate(state = "Assam") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Bihar
```{r bihar}
bihar <- read_clean(85,5)

# Look at the data
kable(bihar) %>% kable_styling()

long_bihar <- make_long(bihar)

ggplot(long_bihar, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(), width = 0.5) + facet_grid(Polling.agency ~ reorder(Date.published, rev(date))) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats") + ggtitle("Bihar (40)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_bihar, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise(average_seats = round(mean(seats))) %>% mutate(state = "Bihar") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Chhattisgarh
```{r chhattisgarh}
chhattisgarh <- read_clean(93,6)

# Look at the data
kable(chhattisgarh) %>% kable_styling()

long_chhattisgarh <- make_long(chhattisgarh)

ggplot(long_chhattisgarh, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(), width = 0.5) + facet_grid(Polling.agency ~ reorder(Date.published, rev(date)) ) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats") + scale_y_continuous(breaks = seq(0,8,2))+ ggtitle("Chhattisgarh (11)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_chhattisgarh, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Chhattisgarh") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Goa
```{r goa}
goa <- read_clean(102,2)

# Look at the data
kable(goa) %>% kable_styling()

long_goa <- make_long(goa)

ggplot(long_goa, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~Polling.agency) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats") + scale_y_continuous(breaks = seq(0,2,1)) +  ggtitle("Goa (2)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_goa, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Goa") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Gujarat
```{r gujarat}
gujarat <- read_clean(107,3)

# Look at the data
kable(gujarat) %>% kable_styling()

long_gujarat <- make_long(gujarat)

ggplot(long_gujarat, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_grid(Polling.agency ~ reorder(Date.published, rev(date))) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats") +  ggtitle("Gujarat (26)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_gujarat, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Gujarat") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Haryana

```{r haryana}
haryana <- read_clean(113,3)

# Look at the data
kable(haryana) %>% kable_styling()

long_haryana <- make_long(haryana)

ggplot(long_haryana, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~Polling.agency) + scale_fill_manual(values = c('#00FF00', '#FFA500', '#228B22'), name = "Party", labels = c("INLD","NDA", "UPA")) + labs(x = "", y = "Number of seats") + ggtitle("Haryana(10)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_haryana, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Haryana") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Himachal Pradesh

## Haryana
```{r hp}
hp <- read_clean(119,2)

# Look at the data
kable(hp) %>% kable_styling()

long_hp <- make_long(hp)

ggplot(long_hp, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~Polling.agency) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats") + ggtitle("Himachal Pradesh (4)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_hp, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Himachal Pradesh") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Jammu and Kashmir
```{r jk}
jk <- read_clean(124,3)

# Look at the data
kable(jk) %>% kable_styling()

long_jk <- make_long(jk)

ggplot(long_jk, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~Polling.agency) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats") +  ggtitle("Jammu and Kashmir (6)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_jk, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Jammu & Kashmir") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Jharkhand
```{r jharkhand}
jharkhand <- read_clean(130,2)

# Look at the data
kable(jharkhand) %>% kable_styling()

long_jharkhand <- make_long(jharkhand)

ggplot(long_jharkhand, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~Polling.agency) + scale_fill_manual(values = c('#FFD700','#FFA500', '#228B22'), name = "Party", labels = c("JVM","NDA", "UPA")) + labs(x = "", y = "Number of seats") +  ggtitle("Jharkhand (14)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_jharkhand, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Jharkhand") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Karnataka
```{r karnataka}
karnataka <- read_clean(135,2)

# Look at the data
kable(karnataka) %>% kable_styling()

long_karnataka <- make_long(karnataka)

ggplot(long_karnataka, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~Polling.agency) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats") +  ggtitle("Karnataka (28)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_karnataka, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Karnataka") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Kerala
```{r kerala}
kerala <- read_clean(140,6)

# Look at the data
kable(kerala) %>% kable_styling()

long_kerala <- make_long(kerala)

ggplot(long_kerala, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~Polling.agency) + scale_fill_manual(values = c('#DC143C','#FFA500', '#6B8E23'), name = "Party", labels = c("LDF","NDA", "UDF")) + labs(x = "", y = "Number of seats") +  ggtitle("Kerala (20)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_kerala, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Kerala") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Madhya pradesh
```{r mp}
mp <- read_clean(149,7)

# Look at the data
kable(mp) %>% kable_styling()

long_mp <- make_long(mp)

ggplot(long_mp, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~Polling.agency) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats")  +  ggtitle("Madhya Pradesh (29)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_mp, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Madhya Pradesh") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Maharashtra
```{r maha}
maha <- read_clean(159,6)

# Look at the data
kable(maha) %>% kable_styling()

long_maha <- make_long(maha)

ggplot(long_maha, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_grid(Polling.agency ~ reorder(Date.published, rev(date))) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats")  +  ggtitle("Maharashtra (48)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_maha, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Maharashtra") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## North East
```{r ne}
ne <- read_clean(168,21)
ne <- ne[c(1,5,9,13,17,21),]
ne$state <- c("Manipur", "Mizoram","Nagaland", "Sikkim", "Arunanchal Pradesh", "Tripura")

# Look at the data
kable(ne) %>% kable_styling()

long_ne <- gather(ne, key = "party", value = "seats", -c(Date.published, Polling.agency, state, Lead))
long_ne$seats <- as.integer(long_ne$seats)

ggplot(long_ne, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~ state) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats") +  scale_y_continuous(breaks = seq(0,2,1)) + ggtitle("Arunanchal Pradesh (2); Manipur (2); Mizoram (1); Nagaland (1); Sikkim (1); Tripura (2)")

meanPoll <- data.frame(party = (c("NONE", "NDA","NDA","NDA","UPA","NDA")), average_seats = as.numeric(str_replace(ne$Lead,ne$Lead[1],"0")), state = ne$state, stringsAsFactors = F)

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Odisha
```{r odisha}
odisha <- read_clean(192,4)

# Look at the data
kable(odisha) %>% kable_styling()

long_odisha <- make_long(odisha)
long_odisha$date <- parse_date_time(long_odisha$Date.published,orders = c("bdy", "bY"))

ggplot(long_odisha, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_grid(Polling.agency ~ reorder(Date.published, rev(date))) + scale_fill_manual(values = c('#6B8E23', '#FFA500', '#228B22'), name = "Party", labels = c("BJD", "NDA", "UPA")) + labs(x = "", y = "Number of seats") +  ggtitle("Odisha (21)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_odisha, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Odisha") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Punjab
```{r punjab}
punjab <- read_clean(199,5)

# Look at the data
kable(punjab) %>% kable_styling()

long_punjab <- make_long(punjab)
long_punjab$date <- parse_date_time(long_punjab$Date.published,orders = c("bdy", "bY"))

ggplot(long_punjab, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_grid(Polling.agency ~ reorder(Date.published, rev(date))) + scale_fill_manual(values = c('#1E90FF', '#FFA500', '#228B22'), name = "Party", labels = c("AAP", "NDA", "UPA")) + labs(x = "", y = "Number of seats") +  scale_y_continuous(breaks = seq(0,12,4)) + ggtitle("Punjab (13)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_punjab, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Punjab") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Rajashthan
```{r rajasthan}
rajasthan <- read_clean(207,7)

# Look at the data
kable(rajasthan) %>% kable_styling()

long_rajasthan <- make_long(rajasthan)
long_rajasthan$date <- parse_date_time(long_rajasthan$Date.published,orders = c("bdy", "bY"))

ggplot(long_rajasthan, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_grid(Polling.agency ~ reorder(Date.published, (date))) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats")  +  scale_y_continuous(breaks = seq(0,25,10)) + ggtitle("Rajasthan (25)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_rajasthan, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Rajasthan") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Tamil Nadu
```{r tn}
tn <- read_clean(217,3)

# Look at the data
kable(tn) %>% kable_styling()

long_tn <- make_long(tn)
long_tn$date <- parse_date_time(long_tn$Date.published,orders = c("bdy", "bY"))

ggplot(long_tn, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~ Polling.agency) + scale_fill_manual(values = c('#DC143C', '#FFA500', '#228B22'), name = "Party", labels = c("AIADMK", "NDA", "UPA")) + labs(x = "", y = "Number of seats") +  ggtitle("Tamil Nadu (39)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_tn, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Tamil Nadu") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Telangana
```{r telangana}
telangana <- read_clean(223,4)

# Look at the data
kable(telangana) %>% kable_styling()

long_telangana <- make_long(telangana)
long_telangana$date <- parse_date_time(long_telangana$Date.published,orders = c("bdy", "bY"))

ggplot(long_telangana, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_grid(Polling.agency ~ reorder(Date.published, rev(date))) + scale_fill_manual(values = c('#6B8E23','#FFA500', '#FF00FF','#228B22'), name = "Party", labels = c("AIMIM", "BJP","TRS","UPA")) + labs(x = "", y = "Number of seats") +  ggtitle("Telangana (17)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_telangana, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Telangana") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Uttar pardesh
```{r up}
up <- read_clean(230,10)

# Look at the data
kable(up) %>% kable_styling()

long_up <- make_long(up)
long_up$date <- parse_date_time(long_up$Date.published,orders = c("bdy", "bY"))

ggplot(long_up, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_grid(Polling.agency ~ reorder(Date.published, rev(date))) + scale_fill_manual(values = c('#FFA500', '#556B2F','#228B22'), name = "Party", labels = c("NDA","SP-BSP", "UPA")) + labs(x = "", y = "Number of seats") +  scale_y_continuous(breaks = seq(0,80,20)) + ggtitle("Uttar Pradesh (80)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_up, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Uttar Pradesh") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Uttarakhand
```{r uttarakhand}
uttarakhand <- read_clean(243,2)

# Look at the data
kable(uttarakhand) %>% kable_styling()

long_uttarakhand <- make_long(uttarakhand)
long_uttarakhand$date <- parse_date_time(long_uttarakhand$Date.published,orders = c("bdy", "bY"))

ggplot(long_uttarakhand, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~ Polling.agency) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats")  +  scale_y_continuous(breaks = seq(0,5,2)) + ggtitle("Uttarakhand (5)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_uttarakhand, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "Uttarakhand") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## West bengal
```{r westbengal}
westbengal <- read_clean(248,4)

# Look at the data
kable(westbengal) %>% kable_styling()

long_westbengal <- make_long(westbengal)
long_westbengal$date <- parse_date_time(long_westbengal$Date.published,orders = c("bdy", "bY"))

ggplot(long_westbengal, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_grid(Polling.agency ~ reorder(Date.published, rev(date))) + scale_fill_manual(values = c('#7CFC00','#8B0000','#FFA500','#228B22'), name = "Party", labels = c("AITC","LF","NDA","UPA")) + labs(x = "", y = "Number of seats") +  ggtitle("West Bengal (42)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_westbengal, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise( average_seats = round(mean(seats))) %>% mutate(state = "West Bengal") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Union territories
```{r ut}
ut <- read_clean(255,18)
ut <- ut[c(1,5,9,13,18),]
ut$ut <- c("Andaman & Nicobar Island", "Chandigarh","Dadara & Nagar Haveli", "Lakshadweep", "Puducherry")

# Look at the data
kable(ut) %>% kable_styling()

long_ut <- gather(ut, key = "party", value = "seats", -c(Date.published, Polling.agency,ut, Lead))
long_ut$seats <- as.integer(long_ut$seats)

ggplot(long_ut, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_wrap(~ ut) + scale_fill_manual(values = c('#FFA500', '#228B22'), name = "Party", labels = c("NDA", "UPA")) + labs(x = "", y = "Number of seats") +  scale_y_continuous(breaks = seq(0,1,1)) + ggtitle("Andaman & Nicobar Island (1); Chandigarh (1); Dadara & Nagar Haveli (1); Lakshadweep (1),  Puducherry (1)")

meanPoll <- data.frame(party = (c("NDA","NDA","NDA","UPA","UPA")), average_seats = c(1,1,1,1,1), state = c("Andaman & Nicobar Island", "Chandigarh", "Dadara & Nagar Havelli", "Lakshadweep", "Puducherry"), stringsAsFactors = F)

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)
```

## Delhi
```{r delhi}
delhi <- read_clean(276,3)

# Look at the data
kable(delhi) %>% kable_styling()

long_delhi <- make_long(delhi)
long_delhi$date <- parse_date_time(long_delhi$Date.published,orders = c("bdy", "bY"))

ggplot(long_delhi, aes(x = party, y = seats, fill = party)) + geom_bar(stat = "identity", position = position_dodge(),width = 0.5) + facet_grid(Polling.agency ~ reorder(Date.published, rev(date))) + scale_fill_manual(values = c('#1E90FF', '#FFA500', '#228B22'), name = "Party", labels = c("AAP", "NDA", "UPA")) + labs(x = "", y = "Number of seats") +  scale_y_continuous(breaks = seq(0,7,2)) + ggtitle("Delhi (7)")

# Calculate the mean poll for 2019
meanPoll <- filter(long_delhi, str_detect(Date.published, "2019")) %>% group_by(party) %>% summarise(average_seats = round(mean(seats))) %>% mutate(state = "NCT of Delhi") %>% slice(which.max(average_seats))

meanPoll2019 <- bind_rows(meanPoll2019, meanPoll)

```

# Indian political map according to opinion polls

```{r map}
india <- rgdal::readOGR("maps/India_map/Admin2.shp")
state_name_id <- data.frame( id = seq(0,35,1), state = as.character(india$ST_NM))

state_poll_info <- left_join(state_name_id, meanPoll2019, by = "state")

fortify_shape <- fortify(india)

merged_data <- merge(fortify_shape,state_poll_info, by = "id", all.x = TRUE)

ggplot(merged_data) + geom_polygon(aes( x = long, y = lat, group = group, fill = party), color = "black", size = 0.5) + coord_map() + scale_fill_manual(values = c('#7CFC00', '#FFA500', '#708090','#808000','#FF00FF','#1E90FF','#228B22','#008B8B'), name = "Party", labels = c("AITC", "NDA", "NONE","SP.BSP","TRS","UDF","UPA","YSRCP")) + labs(x = "Longitude", y = "Latitude") + ggtitle("Political map of India post 2019 elections")
```